Large scale risks from agricultural pesticides in small streams

### 2017-02-02 article.tex
More revisions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Pesticides small streams]{Large scale risks from pesticides in small streams}
% \abbreviations{mo, neon, ra, tu, fw}
\keywords{Monitoring, Neonicotinoid, Risk Assessment, Exposure, Freshwater}

@@

Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to bigger streams \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
For example, the WFD defines small streams with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines small streams with catchment size \textless 10km\textsuperscript{2}. 
Moreover, small streams might particularly be at high risk of pesticide contamination in case of adjacent agricultural areas and given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies was conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}. 
Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant. 
In European law, the Directive 2009/128/EC \citep{European_Union_2009} places an obligation on the EU Member States to adopt National Action Plans (NAP) for the Sustainable Use of Plant Protection Products and the German NAP also addresses the knowledge gap concerning pesticide impact on small streams, specifically including those with catchment size \textless 10km\textsuperscript{2}.

In this study, we compiled and analysed large-scale chemical monitoring data from small streams in Germany. 
First, we analysed the shape of the relationship between pesticide risk, agricultural land use, and catchment size and examined whether related thresholds for pesticide risks can be derived. 
Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010,schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales. 
Finally, we quantified the current risks from pesticides in small streams in Germany and the compounds accountable for the risk.

@@

where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
We modelled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
$\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}. 
The degree of smoothness was estimated using restricted maximum likelihood (REML) during the model fitting process \citep{wood_fast_2011}.
The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modelling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to identify regions of statistically significant changes.
All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Pesticides small streams]{Large scale risks from agricultural pesticides in small streams}
% \abbreviations{mo, neon, ra, tu, fw}
\keywords{Monitoring, Neonicotinoid, Risk Assessment, Exposure, Freshwater}

@@

Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to bigger streams \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
For example, the WFD defines small streams with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines small streams with catchment size \textless 10km\textsuperscript{2}. 
Moreover, small streams might particularly be at high risk of pesticide contamination in case of adjacent agricultural areas and given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
Despite their ecological relevance and potentially higher pesticide exposure, a recent review of pesticide studies showed that a disproportionally small fraction of studies was conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}. 
Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant. 
In European law, the Directive 2009/128/EC \citep{European_Union_2009} places an obligation on the EU Member States to adopt National Action Plans (NAP) for the Sustainable Use of Plant Protection Products and the German NAP also addresses the knowledge gap concerning pesticide impact on small streams, specifically including those with catchment size \textless 10km\textsuperscript{2}.

In this study, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany in order to identify drivers and dynamics of pesticide concentrations.
We expect that the landscape is a determinant of measured pesticide concentrations.
Because a major fraction of pesticides is applied to agricultural fields, we hypothesised highest concentrations and possible exceedances of RACs in streams with high proportion of agriculture. 
Moreover, if agriculture is a main source for pesticides in streams, we expect that concentrations drop to zero if there is not agriculture in the catchment. 
Moreover, these relationships may show thresholds that could be to define reference streams without pollution.
Given their low dilution potential we expected that small streams show highest concentrations. 
However, also the timing of sampling may influence measured concentrations:
A sampling directly after a precipitation might show higher concentrations because of run-off.
Furthermore, pesticides are not applied throughout the whole year and we expected highest concentrations during the main growing season.
Finally, we quantified the current risks from pesticides in small streams in Germany and the compounds accountable for the risk.

@@

where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
Because of overdispersion, we modelled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
$\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}. 
The degree of smoothness was estimated using restricted maximum likelihood (REML) during the model fitting process \citep{wood_fast_2011}.
The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modelling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to identify regions of statistically significant changes.
All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.
